\subsection{Algorithm}

\citeauthor{szwarc_novel_2022} applied the concepts explained in \cref{subsec:04:hs} to the OP. \cite{szwarc_novel_2022}

The solution vectors in $HM$ are arrays that store the nodes visited by each path in the order in which they are visited.
An example is depicted in \cref{fig:04:hm}.

\begin{figure}
    \centering
    \begin{tikzpicture}[every node/.style={draw,minimum width=1cm,minimum height=1cm}]
        \matrix [row sep=3mm] (mat) { 
            \node[draw=none] (0) {$0$}; & \node {1}; & \node{3}; & \node{2}; & \node{5}; & \node {6}; \\
            \node[draw=none] {$1$}; & \node {1}; & \node[fill=yellow] {4}; & \node[fill=yellow!20] {3};& \node{6};  \\
            \node[draw=none] {$2$}; & \node {1}; & \node{2}; & \node[fill=yellow] {4}; & \node[fill=yellow!20]{5}; & \node{6}; \\
        };
        \node[draw=none, above=0cm of 0] {$i$};
        \node[draw=none, left=5mm of mat] {$HM$}; 
    \end{tikzpicture}
    \caption{An example harmony memory with paths starting at node $1$ and ending at node $6$.}
    \label{fig:04:hm}
\end{figure}

This results in the solution vectors not being the same size.
This is not an issue, but it might raise the question of how values for a new harmony, that is a new path, are chosen.

\subsubsection{Choosing the next Node}

Of course the path always starts at the start node, in the case of the example in \cref{fig:04:hm}, it is node $1$.
Say the path so far is $1, 3, 4$.
Which node should we pick next?
As described in \cref{subsec:04:hs}, we have three possibilities.

\paragraph{Choosing from $HM$}

We search for the occurrences of the current node $4$ in $HM$ and keep track of the nodes that appear \emph{after} $4$ in $HM$.
In our case, only $3$ and $5$ appear after.
This is the tentative list of candidates for the next node.
From this list we need to remove all nodes that already appeared in the path and also remove all nodes that would not allow travelling to the end node due to violating $T_{max}$.
We then weigh the 
We remove $3$ from the list of candidates since it already appeared in the path.
Let's say $5$ does not violate $T_{max}$.

\info{Finish describing this}

\paragraph{Choosing from $HM$ and Modifying}

\info{Describe this}

\paragraph{Choosing a random Node}

\info{Describe this}

\subsubsection{Creating a new harmony}

